---
title: "Analysis of Wheat for breeding characters"
author: "Deependra Dhakal"
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2:
    toc: yes
    toc_depth: 3
    keep_tex: yes
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage[normalem]{ulem}
- \usepackage{xcolor}
- \usepackage{rotating}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{subfig}
- \usepackage[format=hang,labelfont=bf,margin=0.5cm,justification=centering]{caption}
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
knitr::opts_chunk$set(tidy = TRUE, cache = TRUE, 
                      echo = FALSE, 
                      tidy.opts = list(width.cutoff=50), 
                      eval = TRUE, warning = FALSE, message = FALSE,
                      # fig.show = "hold", 
                      fig.align = "center", fig.width = 5,
                      fig.asp = 0.9, out.width = "0.7\\linewidth")
options(knitr.kable.NA = "", digits = 3)
# require(lmerTest)
require(agricolae)
```

# Phenotypic traits

## Phenotypic data import

```{r load-data}
normal <- readxl::read_xls("./pramoda-thaguna.xls", sheet = "normal") %>% 
  mutate_at(vars(`Genotypes`, `Replication`), as.factor)

early <- readxl::read_xls("./pramoda-thaguna.xls", sheet = "early") %>% 
  mutate_at(vars(`treatment`, `Replication`), as.factor) %>% 
  select(-`plot no`)
```

<!-- ## Analysis of phenotypic data -->

```{r analysis-and-results}
# rcbd model
early_model <- map(c("DF", "DM", "pH", "PL", 
                     "Tiller", "GY/HA", "TW", "sterility",
                     "grain/spike", "BM", "dryM", "HI"),
                   .f = ~ lm(get(.x) ~ treatment+Replication, 
                             data = early))

normal_model <- map(c("DF", "DM", "pH", "PL", 
                      "Tiller", "GY/HA", "TW", "sterility", 
                      "grain/spike", "BM", "HI"),
                    .f = ~ lm(get(.x) ~ Genotypes+Replication, 
                              data = normal))

# mean separation
## normal
map_inner_normal <- function(y)map_dfr(list("Genotypes" 
                                     # "Replication"
                                     ), 
                  function(x)agricolae::duncan.test(y, trt = x) %>% pluck("groups") %>% 
                    rownames_to_column("Treatment"))

two_fac_gmeans_normal <- map(normal_model, map_inner_normal)
two_fac_gmeans_normal <- map(.x = two_fac_gmeans_normal,
                      .f = ~ .x %>% rename("Mean"="get(.x)"))

## early
map_inner_early <- function(y)map_dfr(list("treatment"
                                           # "Replication"
                                           ), 
                  function(x)agricolae::duncan.test(y, trt = x) %>% pluck("groups") %>% 
                    rownames_to_column("Treatment"))

two_fac_gmeans_early <- map(early_model, map_inner_early)
two_fac_gmeans_early <- map(.x = two_fac_gmeans_early,
                      .f = ~ .x %>% rename("Mean"="get(.x)"))

numeric_names_early <- early %>% map_lgl(is.numeric) %>% which() %>% names()
numeric_names_normal <- normal %>% map_lgl(is.numeric) %>% which() %>% names()

early_gmeans_df <- two_fac_gmeans_early %>% 
  purrr::reduce(left_join, by = "Treatment") %>% 
  magrittr::set_colnames(c("Treatment", paste(rep(numeric_names_early, each = 2), c("", "group"))))

normal_gmeans_df <- two_fac_gmeans_normal %>% 
  purrr::reduce(left_join, by = "Treatment") %>% 
  magrittr::set_colnames(c("Treatment", paste(rep(numeric_names_normal, each = 2), c("", "group"))))
  
# cvs of all traits
early_model_cv <- data.frame(trait = numeric_names_early, 
                          cv = early_model %>% 
                            map_dbl(agricolae::cv.model))

normal_model_cv <- data.frame(trait = numeric_names_normal, 
                          cv = normal_model %>% 
                            map_dbl(agricolae::cv.model))

# # write model cvs to csv
normal_model_cv %>% write_csv(path = "normal_cv_results.csv")
early_model_cv %>% write_csv(path = "early_cv_results.csv")

# write a single file to csv
early_gmeans_df %>%
  write_csv(path = "dmrt_means_result_early.csv",
            na = "")

normal_gmeans_df %>%
  write_csv(path = "dmrt_means_result_normal.csv",
            na = "")
```


## Table of means

```{r two-fac-groups-tab1, results='asis'}

walk2(.x = two_fac_gmeans_normal, 
      .y = numeric_names_normal %>% 
        stringr::str_replace_all("_", "\ ") %>% 
        stringr::str_replace_all("%", "\\\\%"),
      .f = ~ knitr::kable(.x, 
                          caption = paste("Treatment means with groups of", .y),
                          format = "latex", longtable = TRUE,
                          digits = 2, booktabs = TRUE, 
                          align = 'l', escape = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c("striped", 
                                                    "hold_position"), 
                                  font_size = 12, position = "center") %>% 
        print())
```


```{r two-fac-groups-tab2, results='asis'}

walk2(.x = two_fac_gmeans_early, 
      .y = numeric_names_early %>% 
        stringr::str_replace_all("_", "\ ") %>% 
        stringr::str_replace_all("%", "\\\\%"),
      .f = ~ knitr::kable(.x, 
                          caption = paste("Treatment means with groups of", .y),
                          format = "latex", longtable = TRUE,
                          digits = 2, booktabs = TRUE, 
                          align = 'l', escape = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c("striped", 
                                                    "hold_position"), 
                                  font_size = 12, position = "center") %>% 
        print())
```


<!-- # Scatterplots with with regression equation -->
<!-- ## Normal -->

```{r scatterplots-with-regeq, eval=FALSE}

normal <- normal %>% 
  rename_all(function(x)str_replace_all(x, "/", "_per_"))
numeric_names_normal <- numeric_names_normal %>% str_replace_all("/", "_per_")

signif_cor <- correlation(normal[, numeric_names_normal]) %>% pluck("pvalue") %>% 
  as.data.frame() %>% 
  # rownames_to_column() %>% 
  map_df(.f = ~ .x < 0.05) %>% 
  map(which) 
  # map_int(length) %>% 
  # sum()
  
my.formula <- y ~ x

# ggplot2 mapper
library(ggpmisc)
map(numeric_names_normal, 
    function(x){
      walk(numeric_names_normal[signif_cor[[x]]],
          function(y){
            gcor <- ggplot(data = normal, 
                   aes(x = get(x), 
                       y = get(y))) +
              geom_smooth(method = "lm", 
                          se=FALSE, color="black", 
                          formula = my.formula) +
              stat_poly_eq(formula = my.formula, 
                           aes(label = paste(..eq.label.., 
                                             ..rr.label.., sep = "~~~")), 
                           parse = TRUE) +
              geom_point() +
              xlab(x) +
              ylab(y) + 
              theme_bw() 
              ggsave(paste("./correlation_scatterplots_normal/", x, y, ".png", sep = ""), plot = gcor)
          })
    })
```


```{r find-combination, eval = FALSE}
signif_combn <- data.frame(V2 = map(signif_cor, ~ numeric_names_normal[.x]) %>% 
             unlist(use.names = FALSE), 
           V1 = map2(.x = numeric_names_normal, 
                     .y = map_int(signif_cor, length), 
                     .f = ~ rep(.x, times = .y)) %>% unlist(use.names = FALSE)) %>% 
  select(V1, V2)

all_combn <- numeric_names_normal %>% 
  combn(m = 2) %>% 
  t() %>% 
  as_tibble()

to_retain_names <- dplyr::semi_join(signif_combn, all_combn) %>% 
  unite(sep = "", remove = TRUE, col = "plotname") %>% 
  pull(plotname) %>% 
  paste(".png", sep = "")

dir_loc <- "./correlation_scatterplots_normal/"
list.files(dir_loc, full.names = TRUE)[list.files(dir_loc) %in% to_retain_names] %>% 
  file.rename(to = paste("./correlation_scatterplots_removed/", 
                         list.files(dir_loc, full.names = FALSE)[list.files(dir_loc) %in% to_retain_names], sep = ""))
```

<!-- ## Early -->

```{r scatterplots-with-regeq2, eval=FALSE}

early <- early %>% 
  rename_all(function(x)str_replace_all(x, "/", "_per_"))
numeric_names_early <- numeric_names_early %>% str_replace_all("/", "_per_")

signif_cor <- correlation(early[, numeric_names_early]) %>% pluck("pvalue") %>% 
  as.data.frame() %>% 
  # rownames_to_column() %>% 
  map_df(.f = ~ .x < 0.05) %>% 
  map(which) 
  # map_int(length) %>% 
  # sum()
  
my.formula <- y ~ x

# ggplot2 mapper
library(ggpmisc)
map(numeric_names_early, 
    function(x){
      walk(numeric_names_early[signif_cor[[x]]],
          function(y){
            gcor <- ggplot(data = early, 
                   aes(x = get(x), 
                       y = get(y))) +
              geom_smooth(method = "lm", 
                          se=FALSE, color="black", 
                          formula = my.formula) +
              stat_poly_eq(formula = my.formula, 
                           aes(label = paste(..eq.label.., 
                                             ..rr.label.., sep = "~~~")), 
                           parse = TRUE) +
              geom_point() +
              xlab(x) +
              ylab(y) + 
              theme_bw() 
              ggsave(paste("./correlation_scatterplots_early/", x, y, ".png", sep = ""), plot = gcor)
          })
    })
```


```{r find-combination2, eval = FALSE}
signif_combn <- data.frame(V2 = map(signif_cor, ~ numeric_names_early[.x]) %>% 
             unlist(use.names = FALSE), 
           V1 = map2(.x = numeric_names_early, 
                     .y = map_int(signif_cor, length), 
                     .f = ~ rep(.x, times = .y)) %>% unlist(use.names = FALSE)) %>% 
  select(V1, V2)

all_combn <- numeric_names_early %>% 
  combn(m = 2) %>% 
  t() %>% 
  as_tibble()

to_retain_names <- dplyr::semi_join(signif_combn, all_combn) %>% 
  unite(sep = "", remove = TRUE, col = "plotname") %>% 
  pull(plotname) %>% 
  paste(".png", sep = "")

dir_loc <- "./correlation_scatterplots_early/"
list.files(dir_loc, full.names = TRUE)[list.files(dir_loc) %in% to_retain_names] %>% 
  file.rename(to = paste("./correlation_scatterplots_removed/", 
                         list.files(dir_loc, full.names = FALSE)[list.files(dir_loc) %in% to_retain_names], sep = ""))
```


# Agroclimatic variables


## Calculations

1. Growing degree days (GDD)
$$
GDD = \frac{T_{max} + T_{min}}{2}-T_b
$$
Where,

$T_b$ = Base temperature = $100^{\circ}C$ for summer crops

$T_b$ = Base temperature = $80^{\circ}C$ for winter crops

2. Helio-thermal unit (HTU)
$$
HTU = GDD \times Duration~of~sunshine~hours
$$

3. Heat use efficiency (HUE)
$$
HUE = \frac{Biomass~yield~(kg~ha^{-1})}{GDD}
$$

4. Helio-thermal unit use efficiency (HTUE)
$$
HTUE = \frac{Biomass~yield~(kg~ha^{-1})}{HTU}
$$


5. Radiation use efficiency (RUE)
$$
RUE = \frac{Biomass~yield~(kg~ha^{-1})}{Radiation~hours}
$$

Notably, GDD can be calculated at both the heading and maturity periods.

### Climate data import

```{r import-climate}
require(lubridate)
parwanipur_201518_1 <- read_csv("./climate/POWER_SinglePoint_Daily_20150101_20190305_027d07N_084d92E_855a69fa.csv", 
                          comment = "#", na = c("-", -99), skip = 23)
parwanipur_201518_1 <- parwanipur_201518_1 %>% 
  mutate(date_rep = date(strptime(paste(YEAR, as.numeric(DOY), sep = " "), format = "%Y %j"))) %>% 
  select(-YEAR, -DOY)

parwanipur_201518_2 <- read_csv("./climate/POWER_SinglePoint_Daily_20150101_20190305_027d07N_084d92E_a98d45f9.csv", 
                                comment = "#", na = c("-", -99), skip = 17)

parwanipur_201518_2 <- parwanipur_201518_2 %>% 
  mutate(date_rep = date(strptime(paste(YEAR, as.numeric(DOY), sep = " "), format = "%Y %j"))) %>% 
  select(-YEAR, -DOY)

parwanipur_201518_3 <- read_csv("./climate/POWER_SinglePoint_Daily_20150101_20190305_027d07N_084d92E_bf2f8301.csv", 
                                commen = "#", na = c("-", -99), skip = 17)

parwanipur_201518_3 <- parwanipur_201518_3 %>% 
  mutate(date_rep = date(strptime(paste(YEAR, as.numeric(DOY), sep = " "), format = "%Y %j"))) %>% 
  select(-YEAR, -DOY)

parwanipur_201518 <- reduce_right(.x = list(parwanipur_201518_1, 
                                            parwanipur_201518_2, parwanipur_201518_3), 
                                  left_join)

# any annoying encodings exist? if yes recode to explicit NA
# map(parwanipur_201518, ~.x == -99) %>% map(which)

parwanipur_201518 <- parwanipur_201518 %>% 
  mutate_all(funs(na_if(., y = -99)))

# # now check to see if they are explicit
# parwanipur_201518$ALLSKY_SFC_SW_DWN %>% is.na() %>% which()
```

### Climate data filter

```{r data-filter}
# early: June 12
# normal: June 21

parwanipur_201518_normal <- parwanipur_201518 %>% 
  filter(date_rep > ymd("2018-06-20")) %>% 
  # ALLSKY_SFC_SW_DWN SRB/FLASHFlux 1/2x1/2 (All Sky Insolation Incident on a Horizontal Surface (MJ/m^2/day)) in solar insolation expressed as daily hours 
  mutate_at("ALLSKY_SFC_SW_DWN", funs(scales::rescale_max(., c(0.5, 12), range(., na.rm = T))))

parwanipur_201518_early <- parwanipur_201518 %>% 
  filter(date_rep > ymd("2018-06-12")) %>% 
  # ALLSKY_SFC_SW_DWN SRB/FLASHFlux 1/2x1/2 (All Sky Insolation Incident on a Horizontal Surface (MJ/m^2/day)) in solar insolation expressed as daily hours 
  mutate_at("ALLSKY_SFC_SW_DWN", funs(scales::rescale_max(., c(0.5, 12), range(., na.rm = T))))

```

### Climatic and phenotypic indices derived

```{r indices-derived}
climatic_indices_early <- early %>% 
  group_by(treatment, Replication) %>%
  summarise(GDD_DF = sum((parwanipur_201518_early$`T2M_MAX`[1:DF]+parwanipur_201518_early$`T2M_MIN`)[1:DF]/2-8), # base temperature for winter and summer crops: 8 and 10
            GDD_DM = sum((parwanipur_201518_early$`T2M_MAX`[1:DM]+parwanipur_201518_early$`T2M_MIN`)[1:DM]/2-8),
            Sunshine_hrs_DF = mean(parwanipur_201518_early$`ALLSKY_SFC_SW_DWN`[1:DF], na.rm = TRUE),
            Sunshine_hrs_DM = mean(parwanipur_201518_early$`ALLSKY_SFC_SW_DWN`[1:DM], na.rm = TRUE),
            HTU_DF = GDD_DF * Sunshine_hrs_DF, 
            HTU_DM = GDD_DM * Sunshine_hrs_DM)

climatic_indices_normal <- normal %>%
  group_by(Genotypes, Replication) %>%
  summarise(GDD_DF = sum((parwanipur_201518_normal$`T2M_MAX`[1:DF]+parwanipur_201518_normal$`T2M_MIN`)[1:DF]/2-8), # base temperature for winter and summer crops: 8 and 10
            GDD_DM = sum((parwanipur_201518_normal$`T2M_MAX`[1:DM]+parwanipur_201518_normal$`T2M_MIN`)[1:DM]/2-8),
            Sunshine_hrs_DF = mean(parwanipur_201518_normal$`ALLSKY_SFC_SW_DWN`[1:DF], na.rm = TRUE),
            Sunshine_hrs_DM = mean(parwanipur_201518_normal$`ALLSKY_SFC_SW_DWN`[1:DM], na.rm = TRUE),
            HTU_DF = GDD_DF * Sunshine_hrs_DF, 
            HTU_DM = GDD_DM * Sunshine_hrs_DM)

# Note DM associated indices are used here
early_indaug <- early %>% 
  left_join(climatic_indices_early) %>% 
  mutate(HUI = `GY/HA`/GDD_DM, 
         HTUE = `GY/HA`/HTU_DM,
         RUE = `GY/HA`/Sunshine_hrs_DM)

normal_indaug <- normal %>% 
  left_join(climatic_indices_normal) %>% 
  mutate(HUI = `GY/HA`/GDD_DM, 
         HTUE = `GY/HA`/HTU_DM,
         RUE = `GY/HA`/Sunshine_hrs_DM)
```


## Table of means

```{r analysis-climate-agroclimate, cache=TRUE}
# rcbd model
early_model_aug <- map(c("GDD_DF", "GDD_DM", "Sunshine_hrs_DF", 
                     "Sunshine_hrs_DM", "HTU_DF", 
                     "HTU_DM", "HUI", "HTUE", "RUE"),
                   .f = ~ lm(get(.x) ~ treatment+Replication, 
                             data = early_indaug))

normal_model_aug <- map(c("GDD_DF", "GDD_DM", "Sunshine_hrs_DF", 
                     "Sunshine_hrs_DM", "HTU_DF", 
                     "HTU_DM", "HUI", "HTUE", "RUE"),
                    .f = ~ lm(get(.x) ~ Genotypes+Replication, 
                              data = normal_indaug))

# mean separation
## normal
map_inner_normal <- function(y)map_dfr(list("Genotypes"), 
                  function(x)agricolae::duncan.test(y, trt = x) %>% pluck("groups") %>% 
                    rownames_to_column("Treatment"))

two_fac_gmeans_normal_aug <- map(normal_model_aug, map_inner_normal)
two_fac_gmeans_normal_aug <- map(.x = two_fac_gmeans_normal_aug,
                      .f = ~ .x %>% rename("Mean"="get(.x)"))

## early
map_inner_early <- function(y)map_dfr(list("treatment"), 
                  function(x)agricolae::duncan.test(y, trt = x) %>% pluck("groups") %>% 
                    rownames_to_column("Treatment"))

two_fac_gmeans_early_aug <- map(early_model_aug, map_inner_early)
two_fac_gmeans_early_aug <- map(.x = two_fac_gmeans_early_aug,
                      .f = ~ .x %>% rename("Mean"="get(.x)"))

numeric_names_early_aug <- early_indaug %>% select(GDD_DF:RUE) %>% colnames()
numeric_names_normal_aug <- normal_indaug %>% select(GDD_DF:RUE) %>% colnames()

early_gmeans_df_aug <- two_fac_gmeans_early_aug %>% 
  purrr::reduce(left_join, by = "Treatment") %>% 
  magrittr::set_colnames(c("Treatment", paste(rep(numeric_names_early_aug, each = 2), c("", "group"))))

normal_gmeans_df_aug <- two_fac_gmeans_normal_aug %>% 
  purrr::reduce(left_join, by = "Treatment") %>% 
  magrittr::set_colnames(c("Treatment", paste(rep(numeric_names_normal_aug, each = 2), c("", "group"))))
  
# cvs of all traits
early_model_cv_aug <- data.frame(trait = numeric_names_early_aug, 
                          cv = early_model_aug %>% 
                            map_dbl(agricolae::cv.model))

normal_model_cv_aug <- data.frame(trait = numeric_names_normal_aug, 
                          cv = normal_model_aug %>% 
                            map_dbl(agricolae::cv.model))

# # # write model cvs to csv
# normal_model_cv_aug %>% write_csv(path = "normal_cv_results_agroclimate.csv")
# early_model_cv_aug %>% write_csv(path = "early_cv_results_agroclimate.csv")
# 
# # write a single file to csv
# early_gmeans_df_aug %>%
#   write_csv(path = "dmrt_means_result_early_agroclimate.csv",
#             na = "")
# 
# normal_gmeans_df_aug %>%
#   write_csv(path = "dmrt_means_result_normal_agroclimate.csv",
#             na = "")
```


```{r two-fac-groups-tab-agroclimate-early, results='asis'}
walk2(.x = two_fac_gmeans_early_aug, 
      .y = numeric_names_early_aug %>% 
        stringr::str_replace_all("_", "\ ") %>% 
        stringr::str_replace_all("%", "\\\\%"),
      .f = ~ knitr::kable(.x, 
                          caption = paste("Treatment means with groups of", .y),
                          format = "latex", longtable = TRUE,
                          digits = 2, booktabs = TRUE, 
                          align = 'l', escape = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c("striped", 
                                                    "hold_position"), 
                                  font_size = 12, position = "center") %>% 
        print())
```


```{r two-fac-groups-tab-agroclimate-normal, results='asis'}
walk2(.x = two_fac_gmeans_normal_aug, 
      .y = numeric_names_normal_aug %>% 
        stringr::str_replace_all("_", "\ ") %>% 
        stringr::str_replace_all("%", "\\\\%"),
      .f = ~ knitr::kable(.x, 
                          caption = paste("Treatment means with groups of", .y),
                          format = "latex", longtable = TRUE,
                          digits = 2, booktabs = TRUE, 
                          align = 'l', escape = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c("striped", 
                                                    "hold_position"), 
                                  font_size = 12, position = "center") %>% 
        print())
```


## Correlation matrix of numeric variables

```{r correlation-annotate}
# This function annotates the correlation coefficients 
# with the stars corresponding to the level of significance
# x is a matrix containing the data
# method : correlation method. "pearson"" or "spearman"" is supported
# removeTriangle : remove upper or lower triangle
# results : "html" or "latex" (default)
# align: a vector of alignment values as is specified in xtable(x, align)

large <- function(x){
  paste0('{\\footnotesize{\\bfseries ', x, '}}')
}

corstars <-function(x, align, caption, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower"),
                    result=c("none", "html", "latex")){
  #Compute correlation matrix
  require(Hmisc)
  x <- as.matrix(x)
  correlation_matrix<-rcorr(x, type=method[1])
  R <- correlation_matrix$r # Matrix of correlation coeficients
  p <- correlation_matrix$P # Matrix of p-value 
  
  ## Define notions for significance levels; spacing is important.
  mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
  
  ## trunctuate the correlation matrix to two decimal
  R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
  
  ## build a new matrix that includes the correlations with their apropriate stars
  Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
  diag(Rnew) <- paste(diag(R), " ", sep="")
  rownames(Rnew) <- colnames(x)
  colnames(Rnew) <- paste(colnames(x), "", sep="")
  
  ## remove upper triangle of correlation matrix
  if(removeTriangle[1]=="upper"){
    Rnew <- as.matrix(Rnew)
    Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
    Rnew <- as.data.frame(Rnew)
  }
  
  ## remove lower triangle of correlation matrix
  else if(removeTriangle[1]=="lower"){
    Rnew <- as.matrix(Rnew)
    Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
    Rnew <- as.data.frame(Rnew)
  }
  
  ## remove last column and return the correlation matrix
  Rnew <- cbind(Rnew[1:length(Rnew)-1])
  if (result[1]=="none") return(Rnew)
  else{
    if(result[1]=="html") print(xtable(Rnew), type="html", comment = FALSE)
    else print(xtable(Rnew, auto = TRUE,  align = align, caption = caption), type="latex", 
               comment = FALSE, size = "\\footnotesize", 
               sanitize.colnames.function = large, 
               booktabs = TRUE, tabular.environment="longtable", 
               floating=FALSE)
  }
} 
```


```{r correlation-normal, results='asis'}
cortable <- corstars(normal_indaug %>% 
                       select_at(which(map_lgl(., is.numeric))))
# # write to csv
# cortable %>% rownames_to_column(var = "traits") %>% write_csv("correlation_normal.csv", na = "")

cortable %>%
  rename_all(function(x)stringr::str_trunc(x, width = 4, side = "c", ellipsis = "")) %>% 
  knitr::kable(format = "latex",
               caption = "Pearson correlation coefficients of normal season phenotypic and agroclimatic traits",
               booktabs = TRUE, align = "l"
               # longtable = TRUE
               ) %>%
  kableExtra::kable_styling(font_size = 14, latex_options = c("striped", "scale_down")) %>%
  # kableExtra::column_spec(1:11, width = "0.11\\\\textwidth") %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::footnote(general = "p < .0001: **** ; p < .001: *** ; p < .01: ** ; p < .05: *",
                       threeparttable = TRUE, escape = FALSE) %>% 
  kableExtra::landscape()

```


```{r correlation-early, results='asis'}
cortable <- corstars(early_indaug %>% 
                       select_at(which(map_lgl(., is.numeric))))
# # write to csv
# cortable %>% rownames_to_column(var = "traits") %>% write_csv("correlation_early.csv", na = "")

cortable %>%
  rename_all(function(x)stringr::str_trunc(x, width = 4, side = "c", ellipsis = "")) %>% 
  knitr::kable(format = "latex",
               caption = "Pearson correlation coefficients of early season phenotypic and agroclimatic traits",
               booktabs = TRUE, align = "l"
               # longtable = TRUE
               ) %>%
  kableExtra::kable_styling(font_size = 14, latex_options = c("striped", "scale_down")) %>%
  # kableExtra::column_spec(1:11, width = "0.11\\\\textwidth") %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::footnote(general = "p < .0001: **** ; p < .001: *** ; p < .01: ** ; p < .05: *",
                       threeparttable = TRUE, escape = FALSE) %>% 
  kableExtra::landscape()
```

## Principal component analysis

### PAM clustering with PCA

```{r pam-pca-clust}
# pam: partitioning around medoids
suppressPackageStartupMessages(require(ggfortify))
ggplot2::autoplot(cluster::pam(normal_indaug %>%
                                 rename_all(function(x)str_replace_all(x, "/", "_per_")) %>% 
                                 group_by(Genotypes) %>%
                                 summarise_if(is.numeric, funs(mean(., na.rm = TRUE))) %>%
                                 select_at(which(map_lgl(., is.numeric))) %>%
                                 scale(), k = 2),
                  frame = TRUE,
                  frame.type = 'norm', # plots ellipse
                  loadings = TRUE, loadings.label = TRUE,
                  loadings.label.size = 5) +
  theme_bw()
```


### PCA biplot

```{r pca-biplot}
# custom created ggplot2 function
# pca biplot (custom created ggplot2 function)
PCbiplot <- function(PC, x="PC1", y="PC2") {
  # PC being a prcomp object
  data <- data.frame(PC$x)
  plot <- ggplot(data, aes(x=get(x), y=get(y))) +
    geom_point() +
    xlab("PC1") +
    ylab("PC2")
  plot <- plot + geom_hline(yintercept = 0, size=.2) + geom_vline(xintercept = 0, size=.2)
  datapc <- data.frame(varnames=rownames(PC$rotation), PC$rotation)
  mult <- min(
    (max(data[,y]) - min(data[,y])/(max(datapc[,y])-min(datapc[,y]))),
    (max(data[,x]) - min(data[,x])/(max(datapc[,x])-min(datapc[,x])))
  )
  datapc <- transform(datapc,
                      v1 = .7 * mult * (get(x)),
                      v2 = .7 * mult * (get(y))
  )
  plot <- plot + coord_equal() +
    geom_text(data=datapc, aes(x=v1, y=v2, label=varnames), size = 5, vjust=1, color="red")
  plot <- plot +
    geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=v2),
                 arrow=arrow(length=unit(0.2,"cm")), alpha=0.75, color="red")
  plot
}

normal_bipcagg <- PCbiplot(prcomp(normal_indaug %>% 
                  rename_all(function(x)str_replace_all(x, "/", "_per_")) %>% 
                  group_by(Genotypes) %>%
                  summarise_if(is.numeric, funs(mean(., na.rm = TRUE))) %>%
                  select_at(which(map_lgl(., is.numeric))) %>%
                  scale())) +
  theme_bw() +
  theme(text = element_text(size = 15))

early_bipcagg <- PCbiplot(prcomp(early_indaug %>% 
                  rename_all(function(x)str_replace_all(x, "/", "_per_")) %>% 
                  group_by(treatment) %>%
                  summarise_if(is.numeric, funs(mean(., na.rm = TRUE))) %>%
                  select_at(which(map_lgl(., is.numeric))) %>%
                  scale())) +
  theme_bw() +
  theme(text = element_text(size = 15))

# ggsave("normal_bipca.png", normal_bipcagg, width = 6, height = 8)
# ggsave("early_bipca.png", early_bipcagg, width = 6, height = 8)
```


## Cluster dendrogram

A cluster dendrogram using average linkage (Unweighted Pair Group Method with Arithmetic Mean; UPGMA) method for proximity calculation showing three discrete clusters (cluster number determined using GAM) is presented in Figure \@ref(fig:dendrogram-viz).

```{r dendrogram-viz, fig.cap="Dendrogram clustering using UPGMA method", fig.keep='all', fig.width=6, out.width='.95\\linewidth', cache=TRUE, eval=TRUE, echo=FALSE, fig.ncol = 2, fig.show="hold", warning=FALSE, message=FALSE}

## for early trial
# crop yield and agro-morphology
hierarchial_clusters <- early_indaug %>% 
  rename_all(function(x)str_replace_all(x, "/", "_per_")) %>% 
  group_by(treatment) %>%
  summarise_if(is.numeric, funs(mean(., na.rm = TRUE))) %>%
  select_at(which(map_lgl(., is.numeric))) %>%
  scale() %>%
  dist() %>%
  hclust("ave")

k <- 3

ggdendro_plants <- hierarchial_clusters %>%
  as.dendrogram() %>%
  dendextend::set("labels_cex", 1.2) %>%
  dendextend::set("labels_col", value = c(3,4,5),
                  k=k) %>%
  dendextend::set("branches_k_color", k = k) %>%
  dendextend::set("branches_lwd", 0.8) %>%
  dendextend::set("leaves_pch", 9) %>%
  dendextend::set("leaves_cex", 0.8) %>%
  dendextend::as.ggdend()

dendr <- ggdendro::dendro_data(hierarchial_clusters, type="rectangle") # convert for ggplot
clust <- cutree(hierarchial_clusters,k=k)  # k clusters
clust.df <- data.frame(label=ggdendro::label(dendr)$label, cluster=as.factor(clust))
dendr[["labels"]] <- merge(dendr[["labels"]],clust.df, by="label")
templabel <- merge(ggdendro::label(ggdendro_plants),
                   enframe(clust, value = "cluster"),
                   by.x = "label", by.y = "name") %>%
  select(-cex, col)
rect <- aggregate(x~cluster,templabel,range)
rect <- data.frame(rect$cluster,rect$x)
ymax <- mean(hierarchial_clusters$height[length(hierarchial_clusters$height)-((2-2):(2-1))])

threenumsum <- function(xy){c(min(xy), (min(xy)+max(xy))/2, max(xy))}

early_dendrogg <- ggplot() +
  geom_segment(data = ggdendro_plants$segments, aes(x=x, y=y, xend=xend, yend=yend)) +
  geom_text(data = ggdendro::label(ggdendro_plants), 
            aes(x=x, y=y, 
                label=paste(" ", unique(early_indaug %>% 
                                          pull("treatment")), 
                            sep = ""), 
                colour = as.factor(col), hjust=0), size=4) +
  coord_flip() +
  scale_y_reverse(expand=c(0.2, 0),
                  breaks = threenumsum(ggdendro_plants$segments$yend),
                  label = c("100%", "50%", "0%")) +
  geom_rect(data=rect, aes(xmin=X1-.3, xmax=X2+.3, ymin=0, ymax=ymax),
            color="red", fill=1:k, linetype = 2, alpha = 0.2)+
  scale_color_manual(labels = c("Cluster 1", "Cluster 2", "Cluster 3"),
                     values = seq_along(1:3)+7) +
  labs(color = "Cluster index\n") +
  ggtitle("") +
  ylab("Similarity") +
  xlab("") +
  theme_classic() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        text = element_text(size = 14))

# ggsave("early_dendrogram_rice.png", early_dendrogg, width = 7, height = 9, units = "in")

## for normal trial
# crop yield and agro-morphology
hierarchial_clusters <- normal_indaug %>% 
  rename_all(function(x)str_replace_all(x, "/", "_per_")) %>% 
  group_by(Genotypes) %>%
  summarise_if(is.numeric, funs(mean(., na.rm = TRUE))) %>%
  select_at(which(map_lgl(., is.numeric))) %>%
  scale() %>%
  dist() %>%
  hclust("ave")

k <- 3

ggdendro_plants <- hierarchial_clusters %>%
  as.dendrogram() %>%
  dendextend::set("labels_cex", 1.2) %>%
  dendextend::set("labels_col", value = c(3,4,5),
                  k=k) %>%
  dendextend::set("branches_k_color", k = k) %>%
  dendextend::set("branches_lwd", 0.8) %>%
  dendextend::set("leaves_pch", 9) %>%
  dendextend::set("leaves_cex", 0.8) %>%
  dendextend::as.ggdend()

dendr <- ggdendro::dendro_data(hierarchial_clusters, type="rectangle") # convert for ggplot
clust <- cutree(hierarchial_clusters,k=k)  # k clusters
clust.df <- data.frame(label=ggdendro::label(dendr)$label, cluster=as.factor(clust))
dendr[["labels"]] <- merge(dendr[["labels"]],clust.df, by="label")
templabel <- merge(ggdendro::label(ggdendro_plants),
                   enframe(clust, value = "cluster"),
                   by.x = "label", by.y = "name") %>%
  select(-cex, col)
rect <- aggregate(x~cluster,templabel,range)
rect <- data.frame(rect$cluster,rect$x)
ymax <- mean(hierarchial_clusters$height[length(hierarchial_clusters$height)-((2-2):(2-1))])

threenumsum <- function(xy){c(min(xy), (min(xy)+max(xy))/2, max(xy))}

normal_dendrogg <- ggplot() +
  geom_segment(data = ggdendro_plants$segments, aes(x=x, y=y, xend=xend, yend=yend)) +
  geom_text(data = ggdendro::label(ggdendro_plants), 
            aes(x=x, y=y, 
                label=paste(" ", unique(normal_indaug %>%
                                          pull("Genotypes")), 
                            sep = ""), 
                colour = as.factor(col), hjust=0), size=4) +
  coord_flip() +
  scale_y_reverse(expand=c(0.2, 0),
                  breaks = threenumsum(ggdendro_plants$segments$yend),
                  label = c("100%", "50%", "0%")) +
  geom_rect(data=rect, aes(xmin=X1-.3, xmax=X2+.3, ymin=0, ymax=ymax),
            color="red", fill=1:k, linetype = 2, alpha = 0.2)+
  scale_color_manual(labels = c("Cluster 1", "Cluster 2", "Cluster 3"),
                     values = seq_along(1:3)+7) +
  labs(color = "Cluster index\n") +
  ggtitle("") +
  ylab("Similarity") +
  xlab("") +
  theme_classic() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        text = element_text(size = 14))
# ggsave("normal_dendrogram_rice.png", normal_dendrogg, width = 7, height = 9, units = "in")
```


## Model summaries

```{r writing-model-summary}
# write model summaries to csvs
normal_model %>% 
  map(~(summary(.x))$"coefficients" %>% 
        as_tibble(rownames = NA) %>% 
        rownames_to_column("Term") %>% 
        add_case(.after = nrow(.))) %>% 
  purrr::walk2(.y = numeric_names_normal %>% str_replace_all("/", "_per_"), 
               .f = ~write_csv(.x, paste0("./normal_model/", .y, ".csv", collapse = ""), na = "", col_names = TRUE))

early_model %>% 
  map(~(summary(.x))$"coefficients" %>% 
        as_tibble(rownames = NA) %>% 
        rownames_to_column("Term") %>% 
        add_case(.after = nrow(.))) %>% 
  purrr::walk2(.y = numeric_names_early %>% str_replace_all("/", "_per_"), 
               .f = ~write_csv(.x, paste0("./early_model/", .y, ".csv", collapse = ""), na = "", col_names = TRUE))

normal_model_aug %>% 
  map(~(summary(.x))$"coefficients" %>% 
        as_tibble(rownames = NA) %>% 
        rownames_to_column("Term") %>% 
        add_case(.after = nrow(.))) %>% 
  purrr::walk2(.y = numeric_names_normal_aug %>% str_replace_all("/", "_per_"), 
               .f = ~write_csv(.x, paste0("./normal_model/", .y, ".csv", collapse = ""), na = "", col_names = TRUE))

early_model_aug %>% 
  map(~(summary(.x))$"coefficients" %>% 
        as_tibble(rownames = NA) %>% 
        rownames_to_column("Term") %>% 
        add_case(.after = nrow(.))) %>% 
  purrr::walk2(.y = numeric_names_early_aug %>% str_replace_all("/", "_per_"), 
               .f = ~write_csv(.x, paste0("./early_model/", .y, ".csv", collapse = ""), na = "", col_names = TRUE))
```


## Heritability estimates

```{r heritability}
early_model %>%
  map(~anova(.x) %>% broom::tidy()) %>% 
  map_dfr(~.x %>% summarise_at("sumsq", function(x)x[1]/sum(x) * 100)) %>% 
  rename(heritability = sumsq) %>% 
  add_column(.before = 1, trait = numeric_names_early) %>% 
  write_csv("./heritability/early_heritability.csv", na = "")

early_model_aug %>%
  map(~anova(.x) %>% broom::tidy()) %>% 
  map_dfr(~.x %>% summarise_at("sumsq", function(x)x[1]/sum(x) * 100)) %>% 
  rename(heritability = sumsq) %>% 
  add_column(.before = 1, trait = numeric_names_early_aug) %>% 
  write_csv("./heritability/early_agroclimate_heritability.csv", na = "")

normal_model %>%
  map(~anova(.x) %>% broom::tidy()) %>% 
  map_dfr(~.x %>% summarise_at("sumsq", function(x)x[1]/sum(x) * 100)) %>% 
  rename(heritability = sumsq) %>% 
  add_column(.before = 1, trait = numeric_names_normal) %>% 
  write_csv("./heritability/normal_heritability.csv", na = "")

normal_model_aug %>%
  map(~anova(.x) %>% broom::tidy()) %>% 
  map_dfr(~.x %>% summarise_at("sumsq", function(x)x[1]/sum(x) * 100)) %>% 
  rename(heritability = sumsq) %>% 
  add_column(.before = 1, trait = numeric_names_normal_aug) %>% 
  write_csv("./heritability/normal_agroclimate_heritability.csv", na = "")

```

## Yield prediction model

```{r yield-prediction}
require(lme4)
early_indaug <- early_indaug %>% 
  rename_all(funs(str_replace_all(., "/", "_per_")))

normal_indaug <- normal_indaug %>% 
  rename_all(funs(str_replace_all(., "/", "_per_")))
  
early_pred_model <- lmer(GY_per_HA ~ TW + (1 | Replication) + treatment + grain_per_spike + Tiller + PL, data = early_indaug)

normal_pred_model <- lmer(GY_per_HA ~ TW + (1 | Replication) + Genotypes + grain_per_spike + Tiller + PL, data = normal_indaug)

anova(early_pred_model)
anova(normal_pred_model)

early_pred_model %>% emmeans::emmeans("treatment") %>% plot()
normal_pred_model %>% emmeans::emmeans("Genotypes") %>% plot()

```

# Write anovas of models
```{r anova-write, eval = FALSE}
walk2(.x = early_model, 
      .y = c("DF", "DM", "pH", "PL", 
             "Tiller", "GY/HA", "TW", "sterility",
             "grain/spike", "BM", "dryM", "HI"),
      ~ write_csv(anova(.x), paste0("./anovas/anova_early_", .y, ".csv", collapse = "")))

walk(.x = normal_model,
     .y = c("DF", "DM", "pH", "PL", 
            "Tiller", "GY/HA", "TW", "sterility", 
            "grain/spike", "BM", "HI"),
     ~ write_csv(anova(.x), paste0("./anovas/anova_normal", .y, ".csv", collapse = "")))

walk2(.x = early_model_aug, 
      .y = c("GDD_DF", "GDD_DM", "Sunshine_hrs_DF", 
             "Sunshine_hrs_DM", "HTU_DF", 
             "HTU_DM", "HUI", "HTUE", "RUE"),
      ~ write_csv(anova(.x), paste0("./anovas/anova_early_aug", .y, ".csv", collapse = "")))

walk2(.x = normal_model_aug, 
      .y = c("GDD_DF", "GDD_DM", "Sunshine_hrs_DF", 
             "Sunshine_hrs_DM", "HTU_DF", 
             "HTU_DM", "HUI", "HTUE", "RUE"),
      ~ write_csv(anova(.x), paste0("./anovas/anova_normal_aug", .y, ".csv", collapse = "")))

```
