---
title: "Comparison of treatments and correlation analysis"
author: "Deependra Dhakal"
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2:
    toc: yes
    toc_depth: 3
    keep_tex: yes
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage[table]{xcolor}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage[normalem]{ulem}
- \usepackage{xcolor}
- \usepackage{rotating}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{subcaption}
- \newcommand{\subfloat}[2][need a sub-caption]{\subcaptionbox{#1}{#2}}
- \usepackage[format=hang,labelfont=bf,margin=0.5cm,justification=centering]{caption}
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(tidy = TRUE, cache = TRUE, 
                      echo = FALSE, 
                      tidy.opts = list(width.cutoff=50), 
                      eval = TRUE, 
                      fig.show = "hold", fig.pos = "H",
                      fig.align = "center", 
                      fig.asp = 0.7, message = FALSE, warning = FALSE)
options(knitr.kable.NA = "", digits = 3)
```


```{r load-data}
# Fetch data
caps_sk <- readxl::read_xlsx("./data.xlsx", sheet = "data_sheet") %>% 
  mutate_at(c("EC", "SANi", "Rep"), as.factor)

# # if changing colnames is a need
# var_nitrogen <- var_nitrogen %>% 
#   magrittr::set_colnames()
```


# PAM clustering with PCA

```{r pam-pca-clust}
# pam: partitioning around medoids
suppressPackageStartupMessages(require(ggfortify))
ggplot2::autoplot(cluster::pam(var_nitrogen %>% 
                                 rename_all(function(x)stringr::str_replace_all(x, " ", "_")) %>% 
                                 select_at(which(map_lgl(., is.numeric))) %>% 
                                 scale(), k = 3),
                  frame = TRUE,
                  frame.type = 'norm', # plots ellipse
                  loadings = TRUE, loadings.label = TRUE,
                  loadings.label.size = 5)
```


# PCA biplot 

```{r pca-biplot}
# custom created ggplot2 function
PCbiplot <- function(PC, x="PC1", y="PC2") {
  # PC being a prcomp object
  data <- data.frame(PC$x)
  plot <- ggplot(data, aes(x=get(x), y=get(y))) +
    geom_point() +
    xlab("PC1") +
    ylab("PC2")
  plot <- plot + geom_hline(yintercept = 0, size=.2) + geom_vline(xintercept = 0, size=.2)
  datapc <- data.frame(varnames=rownames(PC$rotation), PC$rotation)
  mult <- min(
    (max(data[,y]) - min(data[,y])/(max(datapc[,y])-min(datapc[,y]))),
    (max(data[,x]) - min(data[,x])/(max(datapc[,x])-min(datapc[,x])))
  )
  datapc <- transform(datapc,
                      v1 = .7 * mult * (get(x)),
                      v2 = .7 * mult * (get(y))
  )
  plot <- plot + coord_equal() +
    geom_text(data=datapc, aes(x=v1, y=v2, label=varnames), size = 5, vjust=1, color="red")
  plot <- plot +
    geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=v2),
                 arrow=arrow(length=unit(0.2,"cm")), alpha=0.75, color="red")
  plot
}

PCbiplot(prcomp(plant_traits %>% 
                  group_by(gen) %>% 
                  summarise_if(is.numeric, mean) %>% 
                  select_at(which(map_lgl(., is.numeric))) %>% 
                  select(-contains("rep")) %>% 
                  scale())) + 
  theme_bw() +
  theme(text = element_text(size = 15))
```

# Cluster dendrogram

A cluster dendrogram using average linkage (Unweighted Pair Group Method with Arithmetic Mean; UPGMA) method for proximity calculation showing three discrete clusters (cluster number determined using GAM) is presented in Figure \@ref(fig:dendrogram-viz).

```{r dendrogram-viz, fig.cap="Dendrogram clustering using UPGMA method", fig.subcap=c("Crop yield and agro-morphological traits", "Seed fatty acid content traits"),  fig.keep='all', fig.width=6, out.width='.80\\linewidth', cache=TRUE, eval=TRUE, echo=FALSE, fig.ncol=1, fig.show="hold", warning=FALSE, message=FALSE}

# crop yield and agro-morphology
hierarchial_clusters <- var_nitrogen %>% 
  group_by(treatment) %>% 
  summarise_if(is.numeric, mean) %>% 
  select_at(which(map_lgl(., is.numeric))) %>% 
  scale() %>%
  dist(method = "euclidean") %>% 
  hclust("ave")

ggdendro_plants <- hierarchial_clusters %>% 
  as.dendrogram() %>% 
  dendextend::set("labels_cex", 1.2) %>% 
  dendextend::set("labels_col", value = c(3,4), 
                  k=2) %>% 
  dendextend::set("branches_k_color", k = 2) %>% 
  dendextend::set("branches_lwd", 0.8) %>% 
  dendextend::set("leaves_pch", 9) %>%
  dendextend::set("leaves_cex", 0.8) %>%
  dendextend::as.ggdend()

dendr <- ggdendro::dendro_data(hierarchial_clusters, type="rectangle") # convert for ggplot
clust <- cutree(hierarchial_clusters,k=2)  # k clusters
clust.df <- data.frame(label=ggdendro::label(dendr)$label, cluster=as.factor(clust))
dendr[["labels"]] <- merge(dendr[["labels"]],clust.df, by="label")
templabel <- merge(ggdendro::label(ggdendro_plants), 
                   enframe(clust, value = "cluster"), 
                   by.x = "label", by.y = "name") %>% 
  select(-cex, col)
rect <- aggregate(x~cluster,templabel,range)
rect <- data.frame(rect$cluster,rect$x)
ymax <- mean(hierarchial_clusters$height[length(hierarchial_clusters$height)-((2-2):(2-1))])

threenumsum <- function(x){c(min(x), (min(x)+max(x))/2, max(x))}

ggplot() +
  geom_segment(data = ggdendro_plants$segments, aes(x=x, y=y, xend=xend, yend=yend)) +
  geom_text(data = ggdendro::label(ggdendro_plants), aes(x=x, y=y, label=paste(" ", (var_nitrogen %>% 
                                                                                 group_by(treatment) %>% 
                                                                                 summarise_if(is.numeric, mean) %>%                                                                                 pull(treatment) %>% levels())[hierarchial_clusters$order], sep = ""), 
                                              colour = as.factor(col), hjust=0), size=4) +
  coord_flip() +
  scale_y_reverse(expand=c(0.2, 0),
                  breaks = threenumsum(ggdendro_plants$segments$yend),
                  label = c("100%", "50%", "0%")) +
  geom_rect(data=rect, aes(xmin=X1-.3, xmax=X2+.3, ymin=0, ymax=ymax), 
            color="red", fill=1:2, linetype = 2, alpha = 0.2)+
  scale_color_manual(labels = c("Cluster 1", "Cluster 2", "Cluster 3"), 
                     values = seq_along(1:2)+7) +
  labs(color = "Cluster index\n") +
  ggtitle("") +
  ylab("Similarity") +
  xlab("") +
  theme_classic() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), 
        text = element_text(size = 14))
```


# Correlation matrix of numeric variables

```{r correlation-annotate}
# This function annotates the correlation coefficients 
# with the stars corresponding to the level of significance
# x is a matrix containing the data
# method : correlation method. "pearson"" or "spearman"" is supported
# removeTriangle : remove upper or lower triangle
# results : "html" or "latex" (default)
# align: a vector of alignment values as is specified in xtable(x, align)

large <- function(x){
  paste0('{\\footnotesize{\\bfseries ', x, '}}')
}

corstars <-function(x, align, caption, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower"),
                    result=c("none", "html", "latex")){
  #Compute correlation matrix
  require(Hmisc)
  x <- as.matrix(x)
  correlation_matrix<-rcorr(x, type=method[1])
  R <- correlation_matrix$r # Matrix of correlation coeficients
  p <- correlation_matrix$P # Matrix of p-value 
  
  ## Define notions for significance levels; spacing is important.
  mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
  
  ## trunctuate the correlation matrix to two decimal
  R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
  
  ## build a new matrix that includes the correlations with their apropriate stars
  Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
  diag(Rnew) <- paste(diag(R), " ", sep="")
  rownames(Rnew) <- colnames(x)
  colnames(Rnew) <- paste(colnames(x), "", sep="")
  
  ## remove upper triangle of correlation matrix
  if(removeTriangle[1]=="upper"){
    Rnew <- as.matrix(Rnew)
    Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
    Rnew <- as.data.frame(Rnew)
  }
  
  ## remove lower triangle of correlation matrix
  else if(removeTriangle[1]=="lower"){
    Rnew <- as.matrix(Rnew)
    Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
    Rnew <- as.data.frame(Rnew)
  }
  
  ## remove last column and return the correlation matrix
  Rnew <- cbind(Rnew[1:length(Rnew)-1])
  if (result[1]=="none") return(Rnew)
  else{
    if(result[1]=="html") print(xtable(Rnew), type="html", comment = FALSE)
    else print(xtable(Rnew, auto = TRUE,  align = align, caption = caption), type="latex", 
               comment = FALSE, size = "\\footnotesize", 
               sanitize.colnames.function = large, 
               booktabs = TRUE, tabular.environment="longtable", 
               floating=FALSE)
  }
} 
```


```{r correlation-postharvest, results='asis'}
# crop yield and agro-morphology
cortable <- corstars(caps_sk %>% 
                       select_at(which(map_lgl(., is.numeric))))

# # write to csv
# cortable %>% 
#   rownames_to_column(var = "traits") %>% 
#   write_csv("correlation_santoshK.csv", na = "")

# html
cortable %>%
  knitr::kable(format = "html",
               caption = "Pearson correlation coefficients of traits",
               booktabs = TRUE, align = "l"
               # longtable = TRUE
               ) %>%
  kableExtra::kable_styling(font_size = 14, bootstrap_options = c("striped", "responsive")) %>%
  # kableExtra::column_spec(1:11, width = "0.11\\\\textwidth") %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::footnote(general = "p < .0001: **** ; p < .001: *** ; p < .01: ** ; p < .05: *",
                       threeparttable = TRUE, escape = FALSE)

cortable %>%
  rename_all(function(x)stringr::str_trunc(x, width = 4, side = "c", ellipsis = "")) %>% 
  knitr::kable(format = "latex",
               caption = "Pearson correlation coefficients of Post harvest traits",
               booktabs = TRUE, align = "l"
               # longtable = TRUE
               ) %>%
  kableExtra::kable_styling(font_size = 14, latex_options = c("striped", "scale_down")) %>%
  # kableExtra::column_spec(1:11, width = "0.11\\\\textwidth") %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::footnote(general = "p < .0001: **** ; p < .001: *** ; p < .01: ** ; p < .05: *",
                       threeparttable = TRUE, escape = FALSE) %>% 
  kableExtra::landscape()

```


# Analyis

```{r analysis-and-results}

# rcbd model
two_fac_model <- map(c("pht 20", "pht 40", "pht 60", "pht 80", "nol 20", "nol 40", "nol 60", 
                       "nol 80", "leaf length 40", "leaf length 60", "leaf length 80", 
                       "leaf breadth40", "leaf breadth 60", "leaf breadth80", "spreading 40", 
                       "spreading 60", "spreading 80", "stalk length", "dm", 
                       # "spreading",
                       "root length", "curd dia", "curd wt", "veg wt", "tss", "ascorbic"),
                     .f = ~ lm(get(.x) ~ variety*nitrogen + 
                                 replication, 
                               data = var_nitrogen))

# mean separation
map_inner <- function(y)map_dfr(list("variety", "nitrogen", 
                                     c("variety", "nitrogen")), 
                  function(x)agricolae::duncan.test(y, trt = x) %>% pluck("groups") %>% 
                    rownames_to_column("Treatment"))
two_fac_gmeans <- map(two_fac_model, map_inner)
two_fac_gmeans <- map(.x = two_fac_gmeans,
                      .f = ~ .x %>% rename("Mean"="get(.x)"))

dmrt_files <- var_nitrogen %>% map_lgl(is.numeric) %>% which() %>% names()
walk2(two_fac_gmeans, paste("./dmrt_results/", dmrt_files, ".csv", sep = ""), 
      .f = ~ write_csv(.x, .y))

# agricolae::SNK.test(ramesh_rcbd_2fac, "Fruit_thinning", console = TRUE)
# agricolae::HSD.test(ramesh_rcbd_2fac, "Fruit_thinning", console = TRUE)
two_fac_gm_names <- map(c("Days_for_Flowering", "Days_for_Fruit_Set", 
            "Days_for_fruit_set_to_Harvest", 
            "Firmness", "pulp_weight", 
            "PWL%_3DAH", "PWL%_5_DAH", 
            "PWL%_7_DAH", "PWL%_9_DAH", "TSS"), ~ paste(.x, c("", "_group"), sep = "")) %>% 
  unlist()

```


# All results summary

```{r two-fac-groups-tab, results='asis'}

walk2(.x = two_fac_gmeans, 
      .y = c("Days for Flowering", "Days for Fruit Set", 
             "Days for fruit set to Harvest", 
             "Firmness", "pulp weight", 
             "PWL\\% 3DAH", "PWL\\% 5 DAH", 
             "PWL\\% 7 DAH", "PWL\\% 9 DAH", "TSS"),
      .f = ~ knitr::kable(.x, 
                          caption = paste("Treatment means with groups of", .y),
                          format = "latex", longtable = TRUE,
                          digits = 2, booktabs = TRUE, 
                          align = 'l', escape = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c("striped", 
                                                    "hold_position"), 
                                  font_size = 12, position = "left") %>% 
        print())
```
